---
title: "hello world"
author:
  - name: andrés
date: 01-14-2021
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  dev.args=list(bg='#fefffc'), 
  code_folding="show code"
  )

# Learn more about creating blogs with Distill at:
# https://rstudio.github.io/distill/blog.html

```

I will be using this site as an exercise in writing / coding. 

_También escribiré algunas cosas en español, pero serán pocas cosas porque estoy viviendo en EE.UU. y necesito practicar mi inglés._

The purpose of this post is to help me figure out how to set up the blog in the first place. ^[I'm using the [__distill__](https://rstudio.github.io/distill/) package to set this up.] 

More interesting stuff will be coming next.

I'll also share some visualizations made with ggplot2.

<aside>
A pie chart is just a stacked bar chart in a polar coordinate system.
</aside>

```{r}
library(ggplot2)

df <- data.frame(type = c("mouth", "body"), size = c(4, 15))

g <- ggplot(df, aes(x = "", y = size, fill = type)) + 
  geom_bar(stat = "identity", show.legend = FALSE)

g
```

```{r}
g + coord_polar(theta = "y")
```

```{r}
g + coord_polar(theta = "y", start = 1) +
    scale_fill_manual(values = c("#FFD104", "#fefffc")) + 
    labs(title = "pacman") +
    theme_void()
```





