% !TEX TS-program = XeLaTeX
%%% ------------------------------------------------------------------------
%%% Requirements
%%% ------------------------------------------------------------------------
%%% xelatex
%%% url.sty
%%% fontspec.sty
%%% fontawesome.sty
%%% hyperrref.sty
%%% xunicode.sty
%%% color.sty
%%% fancyhdr.sty
%%% memoir.cls
%%% ------------------------------------------------------------------------
%%% org-preamble-xelatex.sty
%%% memoir-article-styles.sty
%%%------------------------------------------------------------------------

\documentclass[11pt,article,oneside]{memoir}
\usepackage{url}
\usepackage{fontawesome6}
\usepackage{fancyhdr}
\usepackage{org-preamble-xelatex} 
\fboxsep0.1em

%%%------------------------------------------------------------------------
%%% META
%%%------------------------------------------------------------------------

\def\myauthor{Andrés Castro Araújo}
\def\myaffiliation{Duke University}
\def\myaddress{Sociology Department}
\def\myemail{andres.castro.araujo@duke.edu}
\def\myweb{acastroaraujo.github.io}
\def\mygithub{acastroaraujo}
\def\mybluesky{acastroaraujo.bsky.social}

%%%------------------------------------------------------------------------
%%% DOCUMENT
%%%------------------------------------------------------------------------

\begin{document}

\setromanfont[
  Mapping={tex-text},
  Numbers={OldStyle},
  Ligatures={Common}]{Minion Pro}
\setsansfont[
  Mapping=tex-text,
  Ligatures={Common}, 
  Colour=AA0000]{Unit-Medium}
\setmonofont[
  Mapping=tex-text,
  Scale=MatchLowercase,
  Contextuals={WordInitial,WordFinal},
  Ligatures=TeX
  ]{Fira Code}
\newfontface\scheader[
  SmallCapsFont={Minion Pro},
  SmallCapsFeatures={Letters=SmallCaps}
  ]{Minion Pro}
\newfontface\addressblock[
  Mapping={tex-text}, 
  Numbers={OldStyle},
  Ligatures={Common}
  ]{Minion Pro}
\pagestyle{fancy}
\renewcommand{\headrulewidth}{0pt}
\fancyfoot[R]{\fontsize{8pt}{10pt}\selectfont Updated on August 2025.}

%%%------------------------------------------------------------------------
%%% LOCAL
%%%------------------------------------------------------------------------

\makeatletter
\newcommand*\mysize{%
  \@setfontsize\mysize{6.8}{9.0}%
}
\makeatother
\newcommand{\marginhead}[1]{\marginpar{\textsf{{\footnotesize\vspace{-1em}\flushright #1}}}}
\renewcommand{\labelitemi}{~}
\def\ind{\hangindent=0.75 true cm\hangafter=1 \noindent}
\def\labelitemi{~}
\renewcommand{\labelitemii}{~}

%%%------------------------------------------------------------------------
%%% CONTACT
%%%------------------------------------------------------------------------

\begin{minipage}[t]{2.95in}
 \flushright {\small 
 \href{http://sociology.duke.edu}{Department of Sociology} \\ 758 Erwin Square Building \\ Duke University \\ \vspace{-0.05in} Durham, \textsc{NC}, 27708} 
\end{minipage}
\hfill
\begin{minipage}[t]{2in}
  \flushright \small  \addressblock
    {\mysize \texttt{\href{\myweb}{\myweb}} \, \faGlobe} \\
    {\mysize \texttt{\href{mailto:\myemail}{\myemail}} \, \faEnvelope} \\
    {\mysize \texttt{\href{http://github.com/acastroaraujo}{\mygithub}} \, \faGithub }  \\ 
    {\mysize \texttt{\href{https://bsky.app/profile/acastroaraujo.bsky.social}{\mybluesky}}\, \faBluesky} \\
\end{minipage}

\bigskip

%%%------------------------------------------------------------------------
%%% EDUCATION
%%%------------------------------------------------------------------------

%% Name 
\noindent{\large\scheader \textsc{ANDRÉS CASTRO ARAÚJO}}
\reversemarginpar

\bigskip

%% Education

\marginhead{\sffamily {{\vskip -0.23em} education}}

\ind \makebox[1cm][l]{Ph.D.} Sociology, Duke University, \emph{In progress.} \par

\ind
{%
  \parshape 1 0.43in \dimexpr\linewidth-0.43in\relax
  \footnotesize
  Dissertation Committee: Kieran Healy, Martin Ruef, Stephen Vaisey,
  \& Craig Rawlings.
  \normalsize\par
}

\ind \makebox[1cm][l]{M.A.} Sociology, Duke University, 2023. \par

\ind \makebox[1cm][l]{M.A.} Quantitative Methods in the Social Sciences, Columbia University, 2019. \par

\ind \makebox[1cm][l]{B.A.} Sociology, Universidad del Rosario, 2016. \par

\ind \makebox[1cm][l]{B.A.} Economics, Universidad del Rosario, 2015. \par

\bigskip

\marginhead{\sffamily {\vskip 0.35em} research \newline interests}
\medskip

\noindent Organizations, Law, Automation

\bigskip

%%%------------------------------------------------------------------------
%%% JOURNAL ARTICLES
%%%------------------------------------------------------------------------
 
\marginhead{\sffamily {\vskip 0.35em} publications}
\medskip

\ind \emph{Peer-Reviewed Articles}\par \vspace{0.075in}

% Paper

\ind \normalsize \colorbox{Gray!15}{Andrés Castro Araújo} \& Nicolás Restrepo Ochoa. 2025. ``How to Make a Functionalist Argument." \emph{Sociological Science}, 12(20), 456–485. \par
\ind \hspace{0.35in} \footnotesize \href{http://dx.doi.org/10.15195/v12.a20}{Print.} \vspace{0.05in}

% Paper

\ind \normalsize Aidan Combs, Graham Tierney, Fatima Alqabandi, Devin Cornell, Gabriel Varela, \colorbox{Gray!15}{Andrés Castro Araújo}, Lisa Argyle, Alexander Volfovsky, and Christopher A. Bail. 2023. ``Perceived gender and political persuasion: a social media field experiment during the 2020 US Democratic presidential primary election." 
\emph{Nature Scientific Reports.} 14051 \par
\ind \hspace{0.35in} \footnotesize \href{https://doi.org/10.1038/s41598-023-39359-0}{Print.} \vspace{0.05in}

\bigskip

\ind \normalsize \emph{Working Papers}\par \vspace{0.075in}

% Paper

\ind \normalsize \colorbox{Gray!15}{Andrés Castro Araújo}. 2025. ``The Authority of Abstractions." \par
\ind \hspace{0.35in} \footnotesize \href{https://osf.io/preprints/osf/ysa7g}{arXiv Print.} \vspace{0.05in}

\bigskip

\ind \normalsize \emph{In Progress}\par \vspace{0.075in}

% Paper

\ind \normalsize \colorbox{Gray!15}{Andrés Castro Araújo}. ``An Ecology of Rulings." \medskip

% Paper

\ind \normalsize \colorbox{Gray!15}{Andrés Castro Araújo}. ``Innovation in Constitutional Decision-Making." \medskip

% Paper

\ind \normalsize \colorbox{Gray!15}{Andrés Castro Araújo} \& Braulio Güémez Graniel. ``Categories in Flux: A Formal Model of Social Classification." \medskip

\bigskip

\ind \normalsize \emph{Other Publications}\par \vspace{0.075in}

% Paper

\ind \normalsize \colorbox{Gray!15}{Andrés Castro Araújo}. 2019. ``El poder detrás de un número. Visibilidad estadística y desigualdad en Colombia y República Dominicana." In: \emph{¿Cómo pensar la desigualdad desde los derechos humanos?} Siglo XXI Editores \par

\ind \hspace{0.35in} \footnotesize \href{https://www.dejusticia.org/wp-content/uploads/2019/07/Co%CC%81mo-pensar-la-desigualdad-desde-los-derechos-humanos_-v2.pdf}{Print.} 

% Paper

\ind \normalsize Rodrigo Uprimny, Sergio Chaparro H. \& \colorbox{Gray!15}{Andrés Castro Araújo}. 2019. ``Bridging the Gap: The Evolving Doctrine on ESCR and Maximum Available Resources." In: \emph{The Future of Economic and Social Rights.} Cambridge University Press \medskip

\ind \normalsize \colorbox{Gray!15}{Andrés Castro Araújo}, Paul Lagunes, and Xiaoxuan Yang. ``Un alcance de la integridad estatal en América Latina." 2019. \emph{Redes} \par

\ind \hspace{0.35in} \footnotesize \href{https://www.redesarrollo.pe/wp-content/uploads/2020/10/Un-Alcance-de-la-integridad-estatal-en-America-Latina.pdf}{Print.} \vspace{0.05in}

\ind \normalsize Paul Lagunes, Xiaoxuan Yang, and \colorbox{Gray!15}{Andrés Castro Araújo}. ``The State of Corruption in Latin America." 2019. \emph{Baker Institute Report} \par

\ind \hspace{0.35in} \footnotesize \href{https://www.bakerinstitute.org/research/state-corruption-latin-america/}{Print.} \vspace{0.05in}

\bigskip

\normalsize

%%%------------------------------------------------------------------------
%%% PRESENTATIONS
%%%------------------------------------------------------------------------

\marginhead{\sffamily {\vskip 0.35em} conference \newline presentations}
\medskip

\ind ``An Ecology of Rulings"

\ind \hspace{0.35in} \footnotesize \textemdash Annual Meeting of the American Sociological Association, August 2025. \normalsize 

\ind ``Categories In Flux: Membership Change and Compositional Effects"

\ind \hspace{0.35in} \footnotesize \textemdash Annual Meeting of the International Network of Analytical Sociology, June 2025. \normalsize

\ind ``Innovation in Constitutional Decision-Making"

\ind \hspace{0.35in} \footnotesize \textemdash UNC-Duke Student Conference, February 2025. \normalsize

\ind ``Functionalist foundations for problem-solving sociologists."

\ind \hspace{0.35in} \footnotesize \textemdash Annual Meeting of the American Sociological Association, August 2024. \normalsize

\ind ``Proper Functions in Society: Selectionist Arguments in Sociology." 

\ind \hspace{0.35in} \footnotesize \textemdash On Purpose Philosophy Conference, Duke University, May 2024 \normalsize

\ind ``How to make functionalist arguments." 

\ind \hspace{0.35in} \footnotesize \textemdash Annual Meeting of the Southern Sociological Society. April 2024. \normalsize

\bigskip 

%%%------------------------------------------------------------------------
%%% SOFTWARE
%%%------------------------------------------------------------------------

\marginhead{\sffamily {{\vskip 0.35em} software}}
\medskip
\noindent\emph{R Packages \vspace{0.05in}}

\ind \href{https://acastroaraujo.github.io/ccc/}{\texttt{ccc}} \& \href{https://acastroaraujo.github.io/cccLLM/}{\texttt{cccLLM}} 

\ind \hspace{0.35in} \footnotesize \textemdash Contains the datasets created for my dissertation work on the Colombian Constitutional Court. \normalsize

\ind \href{https://acastroaraujo.github.io/protoClassification/}{\texttt{protoClassification}} 

\ind \hspace{0.35in} \footnotesize \textemdash Simulates data from prototype models of classification. \normalsize

\ind \href{https://acastroaraujo.github.io/actsims/}{\texttt{actsims}} 

\ind \hspace{0.35in} \footnotesize \textemdash Fast implementation of computations related to Affect-Control Theory. \normalsize

\bigskip

%%%------------------------------------------------------------------------
%%% TEACHING
%%%------------------------------------------------------------------------

\marginhead{\sffamily {{\vskip 0.35em} teaching}}
\bigskip

\ind Teaching Assistant

\ind \hspace{0.35in} \footnotesize \textemdash Data Analysis and Visualization (undergraduate), Duke University, 2025

\ind \hspace{0.35in} \footnotesize \textemdash Social Theory (undergraduate), Duke University, 2024

\ind \hspace{0.35in} \textemdash Social Statistics (graduate), Duke University, 2023-2024, (\href{https://acastroaraujo.github.io/socStats/}{Part I}, \href{https://acastroaraujo.github.io/socStats2/}{Part II})

\ind \hspace{0.35in} \textemdash Markets and Management (undergraduate), Duke University, 2022

\ind \hspace{0.35in} \textemdash Social Statistics (undergraduate), Duke University, 2022

\ind \hspace{0.35in} \textemdash Summer Institute in Computational Social Science, Duke University, 2021

\normalsize \ind Instructor 

\footnotesize

\ind \hspace{0.35in} \textemdash Economy and Society, Universidad del Rosario, 2017

\ind \hspace{0.35in} \textemdash Introduction to Statistics, Universidad del Rosario, 2016

\normalsize

\bigskip

%%%------------------------------------------------------------------------
%%% GRANTS AND AWARDS
%%%------------------------------------------------------------------------

\marginhead{\sffamily {\vskip 0.35em}grants and \newline awards \newline}
\medskip

\ind Vorsanger-Smith Scholar Award, Duke Sociology Department, 2025.

\ind Research Grant, Duke Sociology Department, 2025.

\ind Summer Research Fellowship, Duke Graduate School, 2020, 2021.

\ind Pasaporte a la Ciencia, awarded by Colciencias (Bogotá), 2017.

\bigskip

%%%------------------------------------------------------------------------
%%% SERVICE
%%%------------------------------------------------------------------------

\marginhead{\sffamily {{\vskip 0.35em} profession}}
\medskip

\ind \makebox[2.5cm][l]{\emph{Membership}} American Sociological Association\par

\ind \makebox[2.5cm][l]{\emph{Reviewer}} American Journal of Sociology, Análisis Político\par

\bigskip

\marginhead{\sffamily {{\vskip 0.35em} professional \newline experience}}
\medskip

\ind \makebox[2.5cm][l]{\emph{Data Scientist}} TBWA/Chiat/Day, 2019-2020

\ind \makebox[2.5cm][l]{\emph{Researcher}} Center for Law, Justice, and Society — Dejusticia, 2015-2017

\ind \makebox[2.5cm][l]{\emph{Translator}} International Sociological Association, 2011-2015

\end{document}
