---
title: "Optimization in R"
description: |
  A few different ways to do optimization
author:
  - name: andrés
categories:
  - R
date: 01-24-2021
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.height = 3, fig.width = 5, fig.align = "center")

library(tidyr)
library(tibble)
```

*Finding the peak of parabola*

$$y = 15 + 10x - 2x^2$$

First we write this as a function.

```{r}
parabola <- function(x) 15 + 10*x - 2*x^2 
```

We can visualize it using `curve()` or `ggplot2`.

```{r}
library(ggplot2)
theme_set(theme_minimal(base_family = "Avenir Next Condensed"))

g <- ggplot() + 
  geom_function(fun = parabola) +
  xlim(0, 5) +
  labs(x = "x", y = "f(x)")

g
```

Then we call `optimize()`, which takes the function as its first argument, the interval as its second, and an optional argument indicating whether or not you are searching for the function's maximum (minimize is the default).

```{r}
out <- optimize(parabola, interval = c(-100, 100), maximum = TRUE)
out

g + geom_vline(xintercept = out$maximum, linetype = "dashed") +
    geom_hline(yintercept = out$objective, linetype = "dashed")
```

There are many more ways of using optimization in R. For example, if you want to find the maximum of a function with many parameters you can use `optim()`.

```{r}
opt <- optim(
  par = 99, ## initial values; use c(...) to do it with many parameters
  fn = parabola, 
  method = "BFGS",
  # this next line is critical: 
  # it tells R to maximize rather than minimize
  control = list(fnscale = -1)
)

opt
```

In statistics we usually try to find the maximum of likelihood functions in order to fit regression models. For example, a simple logistic regression can be fit by doing the following:

```{r, echo=FALSE}
ucb <- tibble::as_tibble(datasets::UCBAdmissions) %>% 
  tidyr::uncount(weights = n) %>% 
  dplyr::rename_all(tolower) %>% 
  dplyr::mutate(admitted = as.integer(admit == "Admitted")) %>% 
  dplyr::select(!admit) %>% 
  dplyr::mutate(across(where(is.character), factor)) %>% 
  as.data.frame() 
```

```{r}
## The dataset, see `help("UCBAdmissions")`
str(ucb)

## the design matrix
X <- model.matrix(admitted ~ gender + dept, data = ucb)
dim(X)

## the outcome variable
y <- ucb$admitted

## the log-likelihood function
log_likelihood <- function(beta, outcome, dmat) {
  
  lp <- dmat %*% beta ## the linear predictor
  p <- plogis(lp)  ## the link function
  
  sum(dbinom(outcome, size = 1, prob = p, log = TRUE))  ## the log-likelihood
}

## maximum likelihood estimate!
opt <- optim(
  par = rep(1, ncol(X)), ## initial values are all 1's
  fn = log_likelihood, method = "BFGS",
  outcome = y, dmat = X,
  # this next line is critical: 
  # it tells R to maximize rather than minimize
  control = list(fnscale = -1)
)

names(opt$par) <- colnames(X)
opt$par
```

We can compare this to the outcome given by R's `glm()` function:

```{r}
fit <- glm(admitted ~ gender + dept, data = ucb, family = binomial(link = "logit"))
coefficients(fit)
```


****

Users of [**Stan**](https://mc-stan.org/) should know that it can be used for optimization as well.

```{r}
library(rstan)
rstan_options(auto_write = TRUE)
writeLines(readLines("simple_parabola.stan"))
opt_stan <- stan_model("simple_parabola.stan")
fit <- optimizing(opt_stan)
fit
```



