---
title: "hello world"
description: ""
author:
  - name: andrés
date: 01-14-2021
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  dev.args=list(bg='#fefffc')
  )

# Learn more about creating blogs with Distill at:
# https://rstudio.github.io/distill/blog.html

```

I will be using this site as an exercise in writing / coding. The purpose of this first post is just to help me figure out how to set up the blog in the first place^[I'm using the [__distill__](https://rstudio.github.io/distill/) package to set this up.]. More interesting stuff will be coming next.

I'll also share some visualizations made with ggplot2. For example, did you know that _a pie chart is just a stacked bar chart in a polar coordinate system_?

```{r}
library(ggplot2)

df <- data.frame(type = c("mouth", "body"), size = c(4, 15))

g <- ggplot(df, aes(x = "", y = size, fill = type)) + 
  geom_bar(stat = "identity", show.legend = FALSE)

g

g + coord_polar(theta = "y")
```

```{r, preview=TRUE}
g + coord_polar(theta = "y", start = 1) +
    scale_fill_manual(values = c("#FFD104", "#fefffc")) + 
    theme_void()
```

<aside>
a note on the margin
</aside>

```{r, preview=FALSE, echo=FALSE, out.width="0%"}
ggplot() + theme_void()
```




